---
title: "Data Pre-processing exercises - IS624 HW2"
output:
  html_document: 
    css: ~/IS624Fall2019/HW3/more/lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
  word_document: default
---
#####Chapter 6 HA 6.2

```{r, load-packages, eval=TRUE, include=FALSE}
library(fpp2)
library(seasonal)
data(plastics)
```
6.2 The plastics data set consists of the monthly sales (in thousands) of product A for a plastics manufacturer for five years.
# plastics Monthly sales of product A for a plastics manufacturer.

a. Plot the time series of sales of product A. Can you identify seasonal fluctuations and/or a trend-cycle? 
There is a seasonal summer peak and winter trough.  The trend is upward in on a linear path.
```{r, eval_plasticsa, eval=TRUE, include=TRUE}
autoplot(plastics) + xlab("Year") + ylab("Sales (in thousands of product A") +
  ggtitle("5-yr monthly sales (in thousands) of product A")
```

b. Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.
```{r, eval_plasticsb, eval=TRUE, include=TRUE}
plastics %>% decompose(type="multiplicative") %>%
  autoplot() + xlab("Year") +
  ggtitle("Classical multiplicative decomposition of plastic sales (in thousands of product A)")
```

c. Do the results support the graphical interpretation from part a?
 Yes.

d. Compute and plot the seasonally adjusted data.
```{r, eval_plasticsd, eval=TRUE, include=TRUE}
plastics %>%
  stl(t.window=13, s.window="periodic", robust=TRUE) %>%
  autoplot()
```

e. Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is    the effect of the outlier?
It accentuates the seasonal data but doesn't impact the trend.
```{r, eval_plasticse, eval=TRUE, include=TRUE}
plastics[50]<-500
plastics %>%
  stl(t.window=13, s.window="periodic", robust=TRUE) %>%
  autoplot()
```

f. Does it make any difference if the outlier is near the end rather than in the middle of the time series?
It appears not to.  Again, only the seasonal data appears to be impacted at the time of the outlier occurance.
```{r, eval_plasticsf, eval=TRUE, include=TRUE}
plastics[10]<-500
plastics %>%
  stl(t.window=13, s.window="periodic", robust=TRUE) %>%
  autoplot()
```

