---
title: "Exponential Smoothing - HW5"
---
##### Chapter 7 HA 7.1, 7.3

```{r, load_libraries, eval=TRUE, include=FALSE}
library(fpp2)
data(pigs)
```

7.1 Consider the **pigs** series — the number of pigs slaughtered in Victoria each month.
The pigs ts contains monthly total number of pigs slaughtered in Victoria, Australia (Jan 1980 – Aug 1995).

a)  Use the **ses()** function in R to find the optimal values of ${\alpha}$ and ${\ell}_{0}$, and generate forecasts for the next four months.

b)  Compute a 95% prediction interval for the first forecast using $\hat{y} {\pm} 1.96s$
  is the standard deviation of the residuals. Compare your interval with the interval produced by R.

```{r, problem_7_1, eval=TRUE, include=TRUE}
# Get timeseries
pigs_ts<-pigs
# Get forecasted estimate parameters from model
round(ses(pigs_ts, h=4)$model$par[1:2],4)
# Forecast out 4 months for visibility on graph
fc_pigs_ses<-ses(pigs_ts, h=4)
# Accuracy of one-step-ahead training errors
round(accuracy(fc_pigs_ses),2)
# see how SES model was fitted
fc_pigs_ses$model
# get 1st 4 months of forecasts
tsCV(pigs_ts,ses,h=4)[1:4,]
# 95% prediction interval for the first forecast
fc_pigs_ses$upper[1, "95%"]
fc_pigs_ses$lower[1, "95%"]
# calculate standard deviation with and without model
s <- sd(fc_pigs_ses$residuals)
s
# s = 10273.69 vs s (estimated) 10308.58
# calculate 95% prediction interval with and without model
fc_pigs_ses$mean[1] + 1.96*s
fc_pigs_ses$mean[1] - 1.96*s
mean(pigs_ts) + 1.96*s
mean(pigs_ts)  - 1.96*s
```


```{r, plot_pigs, eval=TRUE, include=TRUE}
# Note that forecast using ses doesn't have a trend component.
fc_pigs_ses<-ses(pigs_ts, h=100)
autoplot(fc_pigs_ses) + 
    autolayer(fitted(fc_pigs_ses), series="fitted") +
    ylab("Total number of Pigs slaughtered in Victoria, Austrailia") + xlab("Jan 1980 – Aug 1995")
```


7.3 Modify your function from the previous exercise to return the sum of squared errors rather than the forecast of the next observation. Then use the **optim()** function to find the optimal values of ${\alpha}$ and ${\ell}_{0}$.

Do you get the same values as the **ses()** function?
my_ses_func returns an l value that is ~.01% different then the ses function.
##      alpha          l 
##     0.2971 77269.3253
##      alpha          l 
##     0.2971 77260.0561
```{r echo=FALSE, message=FALSE, warning=FALSE, Question3}
# modify ses function to return SSE
my_ses_func <- function(pars=c(alpha, level), y){
  Err <- 0
  SSE <- 0 # calculation varies from 
  for(i in 1:length(y)){
    Err <- y[i] - pars[2]
    SSE <- SSE + Err^2
    pars[2] <- pars[1] * y[i] + (1 - pars[1]) * pars[2]
  }
  return(SSE)
}

# compare ses and my_ses_fun output using the 
opt_ses_pigs <- optim(par = c(alpha = 0.2971, l = pigs_ts[1]), y = pigs_ts, fn = my_ses_func)

round(opt_ses_pigs$par,4)
round(fc_pigs_ses$model$par,4)
```
  
